The value of being indispensable: a cooperative approach to bargaining with a continuum of players

### 2024-01-30 src/paper/theory.tex
Improve structure

\begin{abstract}
    This paper investigates the idea of using concepts random order values, a concept from cooperative game theory, to model bargaining outcomes in games with an indispensable player and a continuum of fringe players.
    I show that this approach proves to be very tractable in a number of settings, and leads to results that are in line with one's intuitive expectations about bargaining outcomes.
    Thus, it can be a useful tool in modelling situations where one does not want to attribute all the bargaining power to one type of market participants, and at the same time would like to avoid having to deal with an explicit multi-stage bargaining game.
    On one hand, this paper provides a taxonomy of the various results found in some applied papers.
    Furthermore, it introduces some new results pertaining to the case when participants have different levels of innate bargaining power, and weighted values are more appropriate.
    The results in this paper is intended to be used as a tool-kit for embedding bargaining outcomes into more complex models.
\end{abstract}

@@

Cooperative games with a finite number of atomic players and a continuum of non-atomic ones are called mixed or oceanic games \parencite{milnor1978values}, and have been studied in the literature of cooperative game theory.
Most papers in this line of research focus on questions such as for which games the Shapley-value exists and is unique \parencite{hart1973values}, and whether various approaches (namely, characterizing the value using a set of axioms, and taking the limit of the value of finite-player games) lead to the same result \parencite{fogelman1980asymptotic}.\footnote{
    \textcite{hart1973values} shows that in the case of mixed games, the usual axioms do not characterize a unique value in general.
    However, \textcite{fogelman1980asymptotic} demonstrates that, for a quite general subset of games, the asymptotic approach leads to one specific element from the set characterized by the axiomatic approach. 
}
My aim with this paper is both more general and more specific in certain aspects.
For one, I am only interested in the asymptotic approach, and withing that, a specific series of finite games.
Furthermore, I restrict my attention to one specific type of games, namely those with an indispensable player, and a continuum of fringe players consisting of a finite number of types.
This in turn allows me to go beyond the usual Shapley-value, and obtain results for a more general set of solution concepts: random order values.
A specific example of those, for which I provide a number of new results, is the weighted value \parencite{shapley1953additive}.
Additionally, due to more structure in the model, I am able to provide more concrete results, investigate some comparative statics, as well.

Despite their theoretical appeal and their tractability, oceanic games saw very little use in the industrial organization, and, more generally, applied economics literature.
The two most prominent examples are \textcite{stole1996intra, stole1996organizational} and \textcite{levy1997individual}, all of which use the Shapley-value to model wage bargaining, and consider the case of divisible labor (continuum of workers).
The first two of these paper also provides microfoundations for the Shapley value and the weighted value in the intra-firm bargaining setting.

In the following sections, I propose an oceanic game with the main feature that the large player is indispensable for creating value.\footnote{
    This assumption will be somewhat relaxed throughout some of the generalizations.
}
Such a model applies to all the settings mentioned above, such as intra-firm bargaining, vertical supply chains, or multi-sided markets.
I demonstrate that the Shapley values of the players in those models are described by very tractable expressions that have nice geometric interpretations.
Additionally, the outcomes are in line with one's intuitive expectations of bargaining power in such situations.
Finally, I demonstrate the usefulness of this approach by building a simple model of TODO.

@@

In this section I examine the case of identical fringe players.
I will first look at the simplest setting (one large player, which is completely indispensable), but in the subsequent sections I will somewhat relax this assumption.
Let us start by defining the cooperative game that describes this situation.

@@

Besides those theoretical directions, the practical applications are no less interesting.
The proposed profit sharing framework could be embedded in more realistic models of various settings.
The example application in this paper is a small example of that, while \textcite{stancsics2023hybrid} provides a more detailed model involving hybrid platforms.
Due to the simplicity of the results, this approach might be a viable replacement of unilateral price setting in cases where one does not wish to ascribe all the bargaining power to one player.

@@

>>>>

\begin{abstract}
    This paper investigates the idea of using concepts random order values, a concept from cooperative game theory, to model bargaining outcomes in games with an indispensable player and a continuum of fringe players.
    I show that this approach proves to be very tractable in a number of settings, and leads to results that are in line with one's intuitive expectations about bargaining outcomes.
    Thus, it can be a useful tool in modelling situations where one does not want to attribute all the bargaining power to one type of market participants, and at the same time would like to avoid having to deal with an explicit multi-stage bargaining game.
    On one hand, this paper provides a taxonomy of the various results found in some applied papers.
    Furthermore, it introduces some new results pertaining to the case of multiple types of fringe players.
    The utility of this approach is demonstrated via a simple model of two-sided markets.
    The results in this paper are intended to be used as a tool-kit for embedding bargaining outcomes into more complex models.
\end{abstract}

@@

Cooperative games with a finite number of atomic players and a continuum of non-atomic ones are called mixed or oceanic games \parencite{milnor1978values}, and have been studied in the literature of cooperative game theory.
Most papers in this line of research focus on questions such as for which games the Shapley-value exists and is unique \parencite{hart1973values}, and whether various approaches (namely, characterizing the value using a set of axioms, and taking the limit of the value of finite-player games) lead to the same result \parencite{fogelman1980asymptotic}.\footnote{
    \textcite{hart1973values} shows that in the case of mixed games, the usual axioms do not characterize a unique value in general.
    However, \textcite{fogelman1980asymptotic} demonstrates that, for a quite general subset of games, the asymptotic approach leads to one specific element from the set characterized by the axiomatic approach. 
}
My aim with this paper is both more general and more specific in certain aspects.
For one, I am only interested in the asymptotic approach, and withing that, a specific series of finite games.
Furthermore, I restrict my attention to one specific type of games, namely those with an indispensable player, and a continuum of fringe players consisting of a finite number of types.
This in turn allows me to go beyond the usual Shapley-value, and obtain results for a more general set of solution concepts: random order values.
In the case of multiple types of fringe players, I also present a number of new results for the Shapley value and the weighted value.
In particular, I derive the value of the different types of fringe players in the multi-sided setting, and demonstrate how they are related to their marginal contributions.
Additionally, due to having more structure in the model, I am able to provide more concrete results, investigate some comparative statics, as well.

Despite their theoretical appeal and their tractability, oceanic games saw very little use in the industrial organization, and, more generally, applied economics literature.
The most prominent examples are \textcite{stole1996intra, stole1996organizational} and \textcite{levy1997individual}, all of which use the Shapley-value to model wage bargaining, and consider the case of divisible labor (continuum of workers).
The first two of these paper also provides microfoundations for the Shapley value and the weighted value in the intra-firm bargaining setting.
To showcase the utility of this approach, I also build a simple model of two-sided markets.
In addition to the aformentioned papers, this model includes two new features: (1) results about the distribution of the value between the various types of fringe players, and (2) using the weighted value in a multi-sided setting.

In the following sections, I propose an oceanic game with the main feature that the large player is indispensable for creating value.\footnote{
    This assumption will be somewhat relaxed throughout some of the generalizations.
}
Such a model applies to all the settings mentioned above, such as intra-firm bargaining, vertical supply chains, or multi-sided markets.
I demonstrate that the random order values of the players in those models are described by very tractable expressions that have nice geometric interpretations.
Additionally, the outcomes are in line with one's intuitive expectations of bargaining power in such situations.
Finally, I demonstrate the usefulness of this approach by building a simple model of two-sided markets, where the various players' shares are determined by their (weighted) Shapley values.

@@

In this section I examine the case of identical fringe players.
Many of these results are known in the literature, but I include them here to build intuition, and highlight the similarities with the multi-sided case.
I will first look at the simplest setting (one large player, which is completely indispensable), but in \ref{sec:extensions} I will somewhat relax this assumption.
Let us start by defining the cooperative game that describes this situation.

@@

Besides those theoretical directions, the practical applications are no less interesting.
The proposed profit sharing framework could be embedded in more realistic models of various settings.
The example application in this paper is a streamlined example of this, while \textcite{stancsics2023hybrid} provides a more detailed model involving hybrid platforms.
Due to the simplicity of the results, this approach might be a viable replacement of unilateral price setting in cases where one does not wish to ascribe all the bargaining power to one player.

@@

\section{Non-indispensable big player(s)}
\label{sec:extensions}

This section demonstrates two ways in which the model can be extended to include non-indispensable big players, while retaining the essential structure and the tractability of the results.
In the first case, there are a number of big players, which are perfect substitutes of each other.
In the second case, there is a single big player, the fringe is able to generate some value on its own, and the big player is only needed to generate the rest.
In both cases, the value function is still a random order value, and thus results from the main text can be applied.
